\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables


% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}
%\makesavenoteenv{tabular}
\usepackage{threeparttable}
\usepackage{siunitx}



% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[PCI '17]{Panhellenic Conference on Informatics}{September 2017}{Larrisa, Greece} 
\acmYear{2017}
\copyrightyear{2017}

\acmPrice{15.00}


\begin{document}
\title{Analyzing Programming Languages' Energy Consumption: An Empirical Study}
\titlenote{Produces the permission block, and
  copyright information}
\subtitle{Extended Abstract}
\subtitlenote{The full version of the author's guide is available as
  \texttt{acmart.pdf} document}


\author{Stefanos Georgiou}
\authornote{Dr.~Trovato insisted his name be first.}
\orcid{1234-5678-9012}
\affiliation{%
  \institution{Athens University of Economics and Business}
  \streetaddress{Patision 76}
  \city{Athens} 
  \state{Greece} 
  \postcode{10434}
}
\email{sgeorgiou@aueb.gr}

\author{Maria Kechagia}
\authornote{The secretary disavows any knowledge of this author's actions.}
\affiliation{%
	\institution{Delft University of Technology}
	\streetaddress{Mekelweg 2}
	\city{Delft} 
	\state{The Netherlands} 
	\postcode{2628 CD}
}
\email{m.kechagia@tudelft.nl}

\author{Diomidis Spinellis}
\authornote{The secretary disavows any knowledge of this author's actions.}
\affiliation{%
  \institution{Athens University of Economics and Business}
  \streetaddress{Patision 76}
  \city{Athens} 
  \state{Greece} 
  \postcode{10434}
}
\email{dds@aueb.gr}


% The default list of authors is too long for headers}
\renewcommand{\shortauthors}{S. Georgiou et al.}


\begin{abstract}
%123456789123456789123456789123456789123456789123456789123456789
\noindent \textbf{Motivation:} Swifting from a traditional 
monolithic approach towards a more cutting-edge such as micro-services 
implies new challenges in terms of energy usage.

%123456789123456789123456789123456789123456789123456789123456789
\noindent \textbf{Goal:} In the preliminary study, we aim is to 
identify energy implications of small, independent tasks developed 
in different programming languages, compiled, semi-compiled, and 
interpreted ones.

%123456789123456789123456789123456789123456789123456789123456789
\noindent \textbf{Method:} To achieve our purpose, we collected, 
refined, compared, and analyzed a number of available tasks from Rosetta 
Code, a publicly open repository for programming chrestomathy. 

%123456789123456789123456789123456789123456789123456789123456789
\noindent \textbf{Results:} Our analysis shows a the majority of 
compiled and semi-compiled programming languages achieve higher 
energy efficiency after applying the optimization options for 
\textit{sorting, object generation, url-encoding} and 
\textit{decoding} tasks. 
Moreover, we show which of the interpreted programming languages 
offers more energy gains than others.

\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
	<ccs2012>
	<concept>
	<concept_id>10010583.10010662.10010674</concept_id>
	<concept_desc>Hardware~Power estimation and optimization</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	<concept>
	<concept_id>10011007.10011006.10011072</concept_id>
	<concept_desc>Software and its engineering~Software libraries and repositories</concept_desc>
	<concept_significance>300</concept_significance>
	</concept>
	<concept>
	<concept_id>10011007.10011074.10011075.10011078</concept_id>
	<concept_desc>Software and its engineering~Software design tradeoffs</concept_desc>
	<concept_significance>300</concept_significance>
	</concept>
	</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Hardware~Power estimation and optimization}
\ccsdesc[300]{Software and its engineering~Software libraries and repositories}
\ccsdesc[300]{Software and its engineering~Software design tradeoffs}


\keywords{GreenIT, Energy Efficiency, Energy Optimization, Programming Languages}


\maketitle

\input{main_body}

\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-bibliography} 

\end{document}
